Nitrogen oxides (NO$_X$) and black carbon (BC) emission factors (EF) of 440 on-road diesel trucks were determined by conducting on-road chasing studies in Beijing and Chongqing, China, in the winter 2010. NO$_X$ and BC EF distributions are reported. The median NO$_X$ EFs for trucks sampled on level roads in Chongqing and Beijing are 40.0 and 47.4 g kg-fuel−1, respectively. The median BC EFs are 1.1 and 0.4 g kg-fuel−1in Chongqing and Beijing, respectively. In addition, a clear downward trend of BC EFs of on-road diesel trucks sampled in Beijing since 2008 is observed. Moreover, Beijing-registered trucks had the lowest BC EFs among the entire sample. These observations appear to reflect the effectiveness of emission standard and fuel quality standard implemented in Beijing (China IV) and nationwide (China III) in reducing BC (and likely overall particulate matter) emission. However, NO$_X$ EFs for Beijing-registered trucks did not show lower value than those from other regions. Unlike black carbon, there is no clear correlation between emission controls and NO$_X$ emissions from the sampled on-road trucks. Further analysis shows that trucks with high BC EFs do not usually have high NO$_X$ EFs, and vice versa, indicating that the current emission standards implemented in Beijing and nationwide have only limited impact on NO$_X$ emissions control. Therefore, effective multi-pollutant control strategies and in-use compliance programs are imperative to reduce the overall emissions from the transportation sector.
